---
title: "Introduction to epiAllele"
author: "Noah Dukler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First create some data and a tree. Pretend that sites are split into two classes, "Enhancer" and "Promoter" and also split by ontology classes "A" and "B".

```{r}
library(epiAllele)
## Set seed for consistency
set.seed(123)
## Simulate random data
species=c("A","B","C")
aData=lapply(as.list(species),function(x) matrix(runif(120),ncol=3))
names(aData)=species

## Create test trees, both one that will fail, and one that will pass
tree=ape::rtree(n = length(species),tip.label = species)

## Create a data frame with labels
siteLabels=data.frame(cre.class=rep(c("Enhancer","Promoter"),each=nrow(aData[[1]])/2),
                      go.class=rep(c("A","B"),times=nrow(aData[[1]])/2))
```
Now create the allele data object:

```{r,echo=FALSE}
ad=alleleData(data=aData,tree=tree,siteInfo = siteLabels)
```

Now create a rate model specifying that genomic regions should be seperated by both cre.class and go.class.

```{r}
rateMod=rateModel(ad,siteLabelCriteria = c("cre.class","go.class"))
```

Note that the model assumes that all different genomic regions have different rate parameters. This can be seen using the plotTree command.
 
Say we want to allow for lineage specific rates. Then we can extract and edge table from the allele data object and add a column called edgeGroup to 

