---
title: "Introduction to epiAllele"
author: "Noah Dukler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First create some data and a tree. Pretend that sites are split into two classes, "Enhancer" and "Promoter" and also split by ontology classes "A" and "B".

```{r}
library(epiAllele)
## Set seed for consistency
set.seed(123)
## Simulate random data
species=c("A","B","C","D","E")
aData=lapply(as.list(species),function(x) matrix(runif(120),ncol=3))
names(aData)=species

## Create test trees, both one that will fail, and one that will pass
tree=ape::rtree(n = length(species),tip.label = species)

## Create a data frame with labels
siteLabels=data.frame(cre.class=rep(c("Enhancer","Promoter"),each=nrow(aData[[1]])/2),
                      go.class=rep(c("A","B"),times=nrow(aData[[1]])/2))
```
Now create the allele data object:

```{r,echo=FALSE}
ad=alleleData(data=aData,tree=tree,siteInfo = siteLabels)
```

Now create a rate model specifying that genomic regions should be seperated by both cre.class and go.class.

```{r}
rateMod=rateModel(ad,rateFormula=formula(~cre.class+go.class))
```

Note that the model assumes that all different genomic regions have different rate parameters. This can be seen using the plotTree command.
 
```{r}
plotTree(obj = rateMod,colorByRate="index")
```
Now we can fit the model and plot the tree with values
```{r}
fittedModel=fit(rateMod)
plotTree(obj = fittedModel$model,colorByRate="value")
```

 
Say we want to allow for lineage specific rates. Then we can extract and edge table from the allele data object and add a column called edgeGroup to 

```{r}
egt=getEdgeTable(ad)
egt[edgeID %in% c("8-4","8-3"),edgeGroup:="e1"]
egt[!edgeID %in% c("8-4","8-3"),edgeGroup:="e2"]
rateMod=rateModel(ad,siteLabelCriteria = c("cre.class","go.class"),lineageTable = egt)
plotTree(rateMod,colorByRate="index")
```

Then say we want to tie rates between two sets of sites for a specific group of edges.

```{r}
rateModTied=tieRates(obj=rateMod,siteLabelA = "Enhancer_A",edgeGroupA = "e1",siteLabelB = "Enhancer_B",edgeGroupB = "e1")
plotTree(obj = rateModTied,colorByRate="index")
```

Then say we want to tie rates between two sets edge groups within a siteLabel

```{r}
rateModelTied2=tieRates(rateMod,siteLabelA = "Enhancer_B",edgeGroupA = "e1",siteLabelB = "Enhancer_B",edgeGroupB = "e2")
plotTree(obj = rateModelTied2,colorByRate="index")
```